Intrinsic alignments from projected tidal fields: non-linear impact on cosmic shear probes

### 2024-04-30 sections/1_introduction.tex
Modifications up to end of Sec. 4

Recent cosmic shear measurements from the Kilo Degree Survey\footnote{KiDS:kids.strw.leidenuniv.nl}, the Dark Energy Survey\footnote{DES:www.darkenergysurvey.org}, and the Hyper Suprime Camera Survey\footnote{HSC:www.naoj.org/Projects/HSC} have established weak gravitational lensing as a competitive probe of cosmology \citep[see. \eg][]{KiDS1000_Asgari, KiDS1000_vdB, KiDS1000_Li, DESY3_Secco, DESY3_Amon, HSCY3_Cl, HSCY3_2pcf}, achieving  percent-level measurements of the structure growth parameter $S_8\equiv\sigma_8\sqrt{\Omega_{\rm m}/0.3}$.
The parameters $\Omega_{\rm m}$ and $\sigma_8$, which respectively describe the abundance and the fluctuations of the matter density fluctuations on scales of $8h^{-1}$ Mpc, are highly degenerate in the lensing signal, and additional data (\eg galaxy clustering data) are required to measure the two individually \citep{Heymans, DES3x2pt, HSC3x2} \niko{fix the refs}.
Despite these highly successful achievements, the current precision of cosmic shear cosmology is mainly limited by the large uncertainty in the intrinsic alignment (IA) of galaxies, a secondary signal that tends to cancel some of the shape correlations produced by lensing \citep[see. \eg][for reviews on IA]{Troxel_IA_review_2015, Kirk_IA_review_2015, Joachimi_IA_review_2015, Kiessling_IA_review_2015}. 
If unaccounted for, the IA can bias by XYZ$\sigma$ the inferred cosmological parameters \citep{BlazekKrause}.
Additionally, using an inaccurate IA model can substantially impair the inference process, as demonstrated by \citet{DESY3_Secco} in their study on DES-Y3 analysis and by \citep{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.
%Furthermore, an incorrect IA model can also cause significant damage to the inference, as shown in \citet{DESY3_Secco} in the context of DES-Y3 analysis, or in \citep{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.

Different physical models have been developed to describe the origin and impact of the IA, including the linear non-linear alignment model \citep[NLA hereafter, ][]{BridleKing2007}, the density-weighted NLA (aka $\delta$-NLA or extended-NLA), and the tidal torque model \citep[][TT hereafter]{BlazekTATT2017}, which respectively assume a linear and quadratic coupling between galaxy ellipticities and the local tidal field. 
The halo model can also describe the IA signal as a function of halo properties such as their mass, concentration, and shapes \citep{FortunaIA}.
A number of observations have sought to place constraints on the parameters from these models \citep[\eg][]{Singh, Sammuroff, Christos, Fortuna, Johnston}.
These observations seem to converge towards a strong colour-dependence: red bright galaxies (elliptical) are generally strongly aligned, with hints of a radial dependence, while blue galaxies (spiral) are consistent with the no-alignment scenario.
%A number of observations have sought to place constraints on the parameters from these models \citep[\eg][]{Singh, Sammuroff, Sammuroff, Christos, Fortuna, Johnston}, which seem to converge towards a strong colour-dependence (red bright galaxies are generally strongly aligned, with hints of a radial dependence), while blue galaxies are consistent with the no-alignment scenario.
The selection effects play a crucial role, making it difficult to generalise these measurements to a different galaxy sample, therefore leaving behind a large uncertainty on the IA parameters.

While most of these methods have been developed to provide prescriptions for modelling IA in lensing  two-point statistics, some have been used to infuse galaxy alignments directly into cosmological simulations, as in \citet{Fluri2019, Tidalator2D, MICE_IA, Borg, Lanzieri}. 
Access to IA-infused numerical simulations is crucial for several applications:
\begin{itemize}
    \item \textit{Validating theoretical models deep in the non-linear regime or utilizing the non-linear galaxy bias model \citep{IA_gal_bias}}.
    \item \textit{Predicting the impact of IA on non-Gaussian lensing probes, for which no models exist \citep{ZuercherXYZ, Tidalator2D}}.
    \item \textit{Testing IA mitigation techniques \citep{IA_selfCalibration}}.
    \item \textit{Exploring the connection between large dark matter haloes and IA \citep{vanAlfen}}.
\end{itemize}
%Having access to such IA-infused numerical simulations is crucial for a number of applications, including the possibility to validate theoretical models deep in the non-linear regime (or using the non-linear galaxy bias model) \citep{IA_gal_bias}, to predict the impact of IA on non-Gaussian lensing probes for which no model exist \citep{ZuercherXYZ, Tidalator2D}, to test IA mitigation techniques \citep{IA_selfCalibration}, or to learn about  the connection between large dark matter haloes and IA \citep{vanAlfen}. 
{\JHD{ (Any other important references?)}}

This paper addresses several of the above-mentioned applications \niko{shall we mention which ones?}.
After reviewing the modelling and measurement of cosmic shear data in Sec. \ref{sec:theory}, we describe our new flexible method to infuse different IA models in cosmological simulations (Sec. \ref{sec:IA_th}). 
Specifically, we infuse the NLA model, the \dNLA, and the TT model on the same underlying lensing simulation. 
In addition, we introduce the \dTT model (which takes into account the fact that galaxies trace dark matter even in the TT model), then proceed to couple the cosmic tidal fields with galaxies taken from halo occupation distributions (HOD) to investigate the impact of realistic non-linear galaxy bias on the IA signal.
These new models are both physically motivated and challenging to describe theoretically as they require perturbation expansions beyond the second order.  
The full numerical implementation is described in Sec. \ref{sec:sims} and validated against theoretical predictions at the level of  two-point shear correlation functions in Sec. \ref{sec:validation}.
Since our infusion method acts on shear galaxy catalogues and on convergence maps, we are in an ideal position to quantify the impact of multiple IA models on different non-Gaussian statistics, which we report in Sec. \ref{sec:HOWLS}, before concluding in Sec. \ref{sec:conclusion}.

>>>>

Recent cosmic shear measurements from the Kilo Degree Survey\footnote{KiDS:kids.strw.leidenuniv.nl}, the Dark Energy Survey\footnote{DES:www.darkenergysurvey.org}, and the Hyper Suprime Camera Survey\footnote{HSC:www.naoj.org/Projects/HSC} have established weak gravitational lensing as a competitive probe of cosmology \citep[see. \eg][]{KiDS1000_Asgari, KiDS1000_vdB, KiDS1000_Li, DESY3_Secco, DESY3_Amon, HSCY3_Cl, HSCY3_2pcf}, achieving  percent-level measurements of the structure growth parameter $S_8\equiv\sigma_8\sqrt{\Omega_{\rm m}/0.3}$.
The parameters $\Omega_{\rm m}$ and $\sigma_8$, which respectively describe the abundance and the amplitude of the fluctuations in the matter density fluctuations on scales of $8h^{-1}$ Mpc, are highly degenerate in the lensing signal, and additional data (\eg galaxy clustering data) are required to measure the two individually from photometric galaxy surveys \citep{KiDS1000_Heymans, DESY3_3x2, HSCY3_3x2}.
Despite these successful achievements, the current precision of cosmic shear cosmology is mainly limited by the large uncertainty in the intrinsic alignment (IA) of galaxies, a secondary signal that tends to cancel some of the shape correlations produced by lensing \citep[see. \eg][for reviews on IA]{Troxel_IA_review_2015, Kirk_IA_review_2015, Joachimi_IA_review_2015, Kiessling_IA_review_2015,Lamman_IA_guide}. 
If unaccounted for, the IA can bias by 4-5$\sigma$ the inferred cosmological parameters \citep{Kirk2012, Krause2016}.
Additionally, using an inaccurate IA model can substantially impair the inference process, as demonstrated by \citet{DESY3_Secco} in their study on DES-Y3 analysis and by \citet{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.
%Furthermore, an incorrect IA model can also cause significant damage to the inference, as shown in \citet{DESY3_Secco} in the context of DES-Y3 analysis, or in \citep{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.

Different physical models have been developed to describe the origin and impact of the IA, including the linear non-linear alignment model \citep[NLA hereafter, ][]{NLA}, the density-weighted NLA (aka $\delta$-NLA or extended-NLA), and the tidal torque model \citep[][TT hereafter]{TATT}, which respectively assume a linear and quadratic coupling between galaxy ellipticities and the local tidal field. 
Alternatively, the halo model can also describe the IA signal as a function of halo properties such as their mass, concentration, and shapes \citep{Fortuna2020}.
A number of observations have sought to place constraints on the parameters from these models \citep[\eg][]{BlueIA, Singh_IA_LOWZ, DESY1_IA_Samuroff, Johnston_IA} {(\it any more recent?)}.
These observations seem to converge towards a strong colour-dependence: red bright galaxies (often elliptical) are generally strongly aligned, with hints of a radial dependence, while blue galaxies (typically spiral) are consistent with the no-alignment scenario.
%A number of observations have sought to place constraints on the parameters from these models \citep[\eg][]{Singh, Sammuroff, Sammuroff, Christos, Fortuna, Johnston}, which seem to converge towards a strong colour-dependence (red bright galaxies are generally strongly aligned, with hints of a radial dependence), while blue galaxies are consistent with the no-alignment scenario.
The selection effects play a crucial role, making it difficult to generalise these measurements to a different galaxy sample, therefore leaving behind a large uncertainty on the IA parameters.

While most theoretical methods have been developed to provide prescriptions for modelling IA in two-point statistics, some have been applied to infuse galaxy alignments directly into cosmological simulations, as in \citet{Fluri2019, Tidalator2D, MICE_IA, Lanzieri2023, vanAlfen2023}. 
Access to such IA-infused numerical simulations is crucial for several applications, including 
validating theoretical models deep in the non-linear regime or utilising non-linear galaxy bias model \citep{IA_gal_bias}, 
 predicting the impact of IA on non-Gaussian lensing probes ({\it i.e.} beyond-2pt statistics), for which no models exist \citep{Zuercher2020a, Tidalator2D},
testing IA mitigation techniques such as self-calibration \citep[][Bera et al in prep.]{SelfCalibrationYao1, SelfCalibrationYao2, SelCalibrationPedersen} or exploring the connection between large dark matter haloes and IA \citep{vanAlfen2023}.
{\JHD{ (Any other important references?)}} These are all important if we are to correctly interpret the cosmological data from photometric redshift surveys.

This paper addresses several of the above-mentioned applications, as we present a novel  pipeline with which we infuse the NLAl, the \dNLA, and the TT model on the same underlying large lensing simulation. 
In addition, we introduce the \dTT model (which takes into account the fact that galaxies trace dark matter even in the TT model), then proceed to couple the cosmic tidal fields with galaxies taken from halo occupation distributions (HOD), thereby probing the impact of realistic non-linear galaxy bias on the IA signal. These new models are both physically motivated and challenging to describe theoretically as they require perturbation expansions beyond the second order.  Existing  beyond-2pt analyses of weak lensing data only model the NLA model \citep{DESY3_Zuercher, HD21}, which is bound to be insufficient with the precision increase provided with the new generation of cosmic shear surveys such as those from the Vera Rubin Observatory \citep{LSST-Design} or {\it Euclid} \citep{RedBook}.

After reviewing the theory and measurement of cosmic shear data in Sec. \ref{sec:theory}, we describe our IA models in Sec. \ref{sec:IA_th}.  
Their full numerical implementation is described in Sec. \ref{sec:sims} and validated against theoretical predictions at the level of  two-point shear correlation functions in Sec. \ref{sec:validation}.
Since our infusion method acts on shear galaxy catalogues and on convergence maps, we are in an ideal position to quantify the impact of multiple IA models on different non-Gaussian statistics, which we report in Sec. \ref{sec:HOWLS}, before concluding in Sec. \ref{sec:conclusion}.

### 2024-04-30 sections/2_cosmic_shear_stats.tex
Modifications up to end of Sec. 4

Cosmic shear data can be analysed using various summary statistics, each with its own advantages and disadvantages.
%Cosmic shear data can be analysed by a number of summary statistics, which all have their pros and cons.
We first review the modelling and measurement methods of shear two-point functions, then present other complementary cosmic shear statistics in the second part of this section.

@@

Shear two-point correlation functions (\gtwopcf hereafter) can be predicted with percent-level precision and are therefore an ideal quantity to validate weak lensing simulations.
In the Limber approximation\footnote{See \citet{Kilbinger17} for a comparison between the Limber approximation and the exact calculations.}, the lensing angular power spectrum $C^{ij}_{\ell}$ (in spherical harmonics space) in tomographic bins $i$ and $j$ is obtained from an integral over the three-dimensional matter power spectrum $P_{\delta}(k, z)$ (in Fourier space, and redshift $z$ dependant):
  \begin{eqnarray}
C_{\ell}^{ij} = \int_0^{\chi_{\rm H}}  \frac{q^i(\chi) \,q^j(\chi)}{\chi^2} \, P_{\delta}\, \bigg(\frac{\ell+1/2}{\chi},z(\chi)\bigg) \ {\rm d}\chi,
\label{eq:C_ell}
\end{eqnarray}
where $\ell$ is a multipole, $\chi_{\rm H}$ is the co-moving distance to the horizon, $c$ is the speed of light, $H_0$ the Hubble parameter, and the lensing kernels $q^{i}$ and $q^j$ are given by:
\begin{eqnarray}
q^{i(j)}(\chi) = \frac{3}{2}\Omega_{\rm m} \, \bigg(\frac{H_0}{c} \bigg)^2 \frac{\chi}{a(\chi)} \int_{\chi}^{\chi_{\rm H}} n^i(\chi')\frac{{\rm d}z}{{\rm d} \chi'}\frac{\chi' - \chi}{\chi'}{\rm d}\chi' \, .
\label{eq:q_lensing}
\end{eqnarray}
In the previous expression, $n^i(z)$ refers to the redshift distribution  in tomographic bins `$i (j)$', while $a(\chi)$ is the scale factor at co-moving distance $\chi$ from the observer.
For more details, consult \citealt{Lamman_IA_guide}.
The matter power spectrum is computed from  {\sc Halofit} \citep{Takahashi2012} in this work, however other public tools provide accurate predictions, including \eg  {\sc HMcode} \citep{HMCode2020}, the {\sc EuclidEmulator} \citep{EuclidEmulator}, the {\sc Bacco} emulator \citep{BACCOEmulator}, or the {\sc MiraTitan} emulator \citep{miraTitan}. 
Predictions for the \gtwopcf are finally computed from Eq. (\ref{eq:C_ell}) as:
\begin{eqnarray}
\xi_{\pm}^{ij}(\theta) = \frac{1}{2\pi} \int_0^{\infty} C_{\ell}^{ij} \, J_{0/4}(\ell \theta) \, \ell \, {\rm d}\ell,
\label{eq:xipm_th}
\end{eqnarray}
where $J_{0/4}(x)$ are Bessel functions of the first kind. 
In this paper, these calculations are carried out by the {\sc CosmoSIS} cosmological inference package \citep{cosmoSIS}.
The redshift distribution $n (z)$ is the source galaxy sample redshift distribution of LSST Year1 \citep{LSST-SRD}:
\begin{eqnarray}
n (z) =  z^2 \: {\rm exp} \left[-\left(\frac{z}{z_0}\right)^\alpha \right]\, ,
\label{eq:nz}
\end{eqnarray}
with pivot redshift $z_0=0.13$ and a power-law index $\alpha = 0.78$.
The distribution is normalised to provide a number density of $n_{\rm eff} = 3.0 $  galaxies per arcmin$^{2}$.
This lower than the expected number density for the first data release ($n_{\rm eff}\sim10$ galaxies arcmin$^{-2}$), but is large enough to validate our methods, and makes the calculation more tractable.
This global $n(z)$ is further split into five equi-populated tomographic bins, and smoothed with a Gaussian filter of width $\sigma = 0.05 \: (1 + z)$ to mimic the photometric selection process, shown by the solid lines in Fig. \ref{fig:Nz}.

The weak lensing signal is measured from the ellipticities $\epsilon_{1/2}$ of simulated or observed galaxies, which, in absence of systematics and secondary signals, are unbiased estimators  of the cosmic shear components $\gamma_{1/2}$.
In particular, the \gtwopcf are estimated as:
\begin{eqnarray}
\widehat{\xi_{\pm}^{ij}}(\theta) = \frac{\sum_{a,b} w_a w_b \left[\epsilon_{a, \rm +}^i\epsilon_{b, \rm +}^j     \pm \epsilon_{a, \times}^i\epsilon_{b, \times}^j    \right]\Delta_{ab}(\theta)}{\sum_{a,b} w_a w_b} \, ,
\label{eq:2PCF_estimator}
\end{eqnarray}
where the sum is over all pairs of galaxies $a, b$ separated by an angular distance $\theta$ on the sky, respectively belonging to the tomographic bins $i$ and $j$.
$w_{a (b)}$ represent the projected correlation functions, while the tangential/cross components of ellipticities are denoted as $\epsilon_{+/\times}$, respectively. 
$\Delta_{ab}(\theta)$ is the binning operator, equal to unity if the angular separation between the two galaxies falls within the $\theta$-bin, and zero otherwise, while the shape weights $w_{a,b}$ are set to unity.
In this work, we construct lensing catalogues from numerical simulations (described in Sec. \ref{sec:sims}), from which we measure $ \widehat{\xi_{\pm}^{ij}}(\theta)$ with {\sc Treecorr} \citep{TreeCorr}.
We set therein the {\sc bin\_slop} parameter to 0.05, then compute the correlations in 20 logarithmically-spaced angular bins with outer edges ranging from $0.5$ to $475.5$ arcmin.

@@

Non-Gaussian statistics have the potential to extract cosmological information stored in the complex phases of the density field, outperforming in that sense the \gtwopcf that can only access information encoded in the complex amplitudes.
No optimal estimator has been identified to-date as being able to capture `all' information, but numerous studies all show a clear gain in constraining power, especially when used in combination with two-point functions \citep{HD21, DESY3_Zuercher, HSC_PeaksSims}.
While some of these act on galaxy catalogues directly, many non-Gaussian statistics are measured on convergence or aperture mass maps, requiring the post-process the galaxy catalogues with a mass-reconstruction algorithm.
To accommodate a variety of cases, we therefore provide both galaxy catalogues and convergence maps, the latter being  produced from the galaxies' observed ellipticities with the standard Kaiser \& Squires inversion technique \citep{KaiserSquires}.
These are spherical {\sc Healpix}\footnote{{\sc Healpix}: http://healpix.sf.net}  maps \citep{healpix} with $N_{\rm side}=4096$ generated over an octant, and includes a baseline $2.0$ arcmin smoothing beam to suppress numerical noise.

It's worth noting that this smoothing scale is relatively small, posing challenges for theoretical modeling of some probes within this regime.
Consequently, modeling often requires direct inference from simulations \citep[see][]{HD21, DESY3_Zuercher, HSCY1_Peaks_sims}.
%Note that this smoothing scale is quite small; some probes are difficult to model theoretically in that regime, in which case the modelling needs to be inferred from simulations directly \citep[as in][]{HD21, DESY3_Zuercher, HSCY1_Peaks_sims}.
In the interest of brevity, this paper does not cover statistics derived from aperture mass maps, though our methods are equally applicable to those estimators.
%We do not discuss statistics based on aperture mass maps in this paper for the sake of conciseness, but all our methods can be applied to those estimators as well.
While tomographic \gtwopcf data include auto-correlation and cross-redshift correlations, certain non-Gaussian statistics extract further information from analyzing triplets, quadruplets, or quintuplets of redshift bins \citep{Martinet21a}.
This work focuses only on auto-tomographic bins, yet our methods and results can be straightforwardly extended to accommodate such analyses.

@@

A number of higher-order statistics work at the level of shear catalogue,  and here we consider the following:
\begin{enumerate}
\item {\it Three-point correlation functions (or $M_{\rm ap}^3(\theta)$)}:  \JHD{ Lucas/Laila, can you summarise the measurements and modelling, and provide references?}
\item {\it Squeezed bispectrum}: \JHD{ Anik, can you summarise the measurements and modelling, and provide references?}
\end{enumerate}

@@

We further consider the following convergence map-based statistics:
\begin{enumerate}
    \item \textit{Convergence three-point correlation function}: \JHD{Alejandro, can you summarise the measurements and modelling, and provide references?}
    \item \textit{Peaks and Minima}: local maxima and minima in the {\sc Healpix} convergence maps are counted in bins of signal-to-noise ratios, $\nu$, computed from the global level of shape noise in the survey. 
    These are amongst the most widely used non-Gaussian lensing statistics, as it is simple to measure yet highly sensitive to non-Gaussian structures.
    Although some theoretical models have been developed to describe the largest peaks \citep[see][]{Shan18, HSCY1_Peaks_th}, we report here the results on a much wider range of statistics, which can only be accurately modelled from numerical simulations.
    \item \textit{Lensing PDF}:  \JHD{Cora/Lina, can you summarise the measurements and modelling, and provide references?}
    \item  \textit{Minkowski functionals}: \JHD{Nisha, can you summarise the measurements and modelling, and provide references?}
\end{enumerate}

These statistics explore various physical scales and nonlinear phenomena, leading to differences in their sensitivity to cosmology and systematic errors.
For instance, a statistical method effective in recovering cosmological information could be significantly influenced by secondary effects like intrinsic alignments (IA), casting doubts on its robustness.
Thus, incorporating IA simulations at the field level is essential for addressing these concerns.

>>>>

Cosmic shear data can be analysed using various summary statistics, each with its own advantages and disadvantages.
%Cosmic shear data can be analysed by a number of summary statistics, which all have their pros and cons.
We start with shear two-point functions, then present other complementary cosmic shear statistics in the second part of this section.

@@

Shear two-point correlation functions (\gtwopcf hereafter) can be predicted with percent-level precision and are therefore an ideal quantity to validate weak lensing simulations.
In the Limber approximation\footnote{See \citet{Kilbinger17} for a comparison between the Limber approximation and the exact calculations.}, the tomographic lensing power spectrum $C^{ij}_{\ell}$, obtained for combinations of redshift bins $i$ and $j$, is calculated from an integral over the three-dimensional matter power spectrum $P_{\delta}(k, z)$ as:
  \begin{eqnarray}
C_{\ell}^{ij} = \int_0^{\chi_{\rm H}}  \frac{q^i(\chi) \,q^j(\chi)}{\chi^2} \, P_{\delta}\, \bigg(\frac{\ell+1/2}{\chi},z(\chi)\bigg) \ {\rm d}\chi,
\label{eq:C_ell}
\end{eqnarray}
where $\ell$ are angular multipoles, $k$ are amplitudes of Fourier modes, $\chi_{\rm H}$ is the comoving distance to the horizon, $c$ is the speed of light, $H_0$ the Hubble parameter, and the lensing kernels $q^{i}$ and $q^j$ are given by:
\begin{eqnarray}
q^{i}(\chi) = \frac{3}{2}\Omega_{\rm m} \, \bigg(\frac{H_0}{c} \bigg)^2 \frac{\chi}{a(\chi)} \int_{\chi}^{\chi_{\rm H}} n^i(\chi')\frac{{\rm d}z}{{\rm d} \chi'}\frac{\chi' - \chi}{\chi'}{\rm d}\chi' \, .
\label{eq:q_lensing}
\end{eqnarray}
In the previous expression, $n^i(z)$ refers to the redshift distribution  in tomographic bins `$i$', while $a(\chi)$ is the scale factor at comoving distance $\chi$ from the observer.
The matter power spectrum is computed from  {\sc Halofit} \citep{Takahashi2012} in this work, however other public tools provide accurate predictions, including \eg  {\sc HMcode} \citep{HMCode2020}, the {\sc EuclidEmulator} \citep{EuclidEmulator}, the {\sc Bacco} emulator \citep{BACCOEmulator}, or the {\sc MiraTitan} emulator \citep{miraTitan}. 
Predictions for the \gtwopcf are finally computed from Eq. (\ref{eq:C_ell}) as:
\begin{eqnarray}
\xi_{\pm}^{ij}(\vartheta) = \frac{1}{2\pi} \int_0^{\infty} C_{\ell}^{ij} \, J_{0/4}(\ell \vartheta) \, \ell \, {\rm d}\ell,
\label{eq:xipm_th}
\end{eqnarray}
where $J_{0/4}(x)$ are Bessel functions of the first kind. 
In this paper, these calculations are carried out by the public\footnote{{\sc CosmoSIS}:https://cosmosis.readthedocs.io/en/latest/ } {\sc CosmoSIS} cosmological inference package \citep{cosmoSIS}.
The redshift distribution $n (z)$ is taken from the LSST Year-1 forecast \citep{LSST-SRD}:
\begin{eqnarray}
n (z) =  z^2 \: {\rm exp} \left[-\left(\frac{z}{z_0}\right)^\alpha \right]\, ,
\label{eq:nz}
\end{eqnarray}
with pivot redshift $z_0=0.13$ and a power-law index $\alpha = 0.78$.
The distribution is normalised to provide a number density of $n_{\rm eff} = 3.0 $  galaxies per arcmin$^{2}$.
This lower than the expected number density for the first data release ($n_{\rm eff}\sim10$ galaxies arcmin$^{-2}$), but is large enough to validate our methods, and makes the calculation more tractable.
This global $n(z)$ is further split into five equi-populated tomographic bins, and smoothed with a Gaussian filter of width $\sigma = 0.05 \: (1 + z)$ to mimic the photometric selection process, shown by the solid lines in Fig. \ref{fig:Nz}.

The weak lensing signal is measured from the ellipticities $\epsilon_{1/2}$ of simulated or observed galaxies, which, in absence of systematics and secondary signals, are unbiased estimators  of the cosmic shear components $\gamma_{1/2}$.
In particular, the \gtwopcf are estimated as:
\begin{eqnarray}
\widehat{\xi_{\pm}^{ij}}(\vartheta) = \frac{\sum_{a,b} w_a w_b \left[\epsilon_{a, \rm +}^i\epsilon_{b, \rm +}^j     \pm \epsilon_{a, \times}^i\epsilon_{b, \times}^j    \right]\Delta_{ab}(\vartheta)}{\sum_{a,b} w_a w_b} \, ,
\label{eq:2PCF_estimator}
\end{eqnarray}
where the sum is over all pairs of galaxies $a, b$ separated by an angular distance $\vartheta$ on the sky, respectively belonging to the tomographic bins $i$ and $j$.
$w_{a (b)}$ represent the weights, describing the precision of the shape measurement, which we ignore in this work, while the tangential/cross components of ellipticities are denoted as $\epsilon_{+/\times}$, respectively. 
$\Delta_{ab}(\vartheta)$ is the binning operator, equal to unity if the angular separation between the two galaxies falls within the $\vartheta$-bin, and zero otherwise.
In this work, we construct lensing catalogues from numerical simulations (described in Sec. \ref{sec:sims}), from which we measure $ \widehat{\xi_{\pm}^{ij}}(\vartheta)$ with {\sc Treecorr} \citep{TreeCorr}.
We set therein the {\sc bin\_slop} parameter to 0.05, then compute the correlations in 20 logarithmically-spaced angular bins with outer edges ranging from $0.5$ to $475.5$ arcmin.

@@

Non-Gaussian statistics have the potential to extract cosmological information stored in the complex phases of the density field, outperforming in that sense the \gtwopcf that can only access information encoded in the complex amplitudes \citep{2002MNRAS.337..488C}.
No optimal estimator has been identified to-date as being able to capture `all' information, but numerous studies all show a clear gain in constraining power, especially when used in combination with two-point functions \citep{Fu2014, Gruen2017, HD21, DESY3_Zuercher, HSCY1_peaks_sims, KiDS1000_Burger, KiDS1000_Map3}.
While some of these act on galaxy catalogues directly, many non-Gaussian statistics are measured on convergence or aperture mass maps, requiring to post-process the galaxy catalogues with a mass-reconstruction algorithm.
To accommodate a variety of cases, we therefore provide both galaxy catalogues and convergence maps, the latter being  produced from the galaxies' observed ellipticities with the standard Kaiser \& Squires (KS)  inversion technique \citep{KaiserSquires}, which relates the shear and convergence to the Newtonian potential. This is achieved by assigning the two ellipticity components of every galaxies on spherical {\sc Healpix}\footnote{{\sc Healpix}: http://healpix.sf.net}  maps \citep{healpix} with $N_{\rm side}=4096$, then converting the shear maps to convergence maps by solving  the KS inversion in spherical harmonic space \citep[][see their equation 10]{Gatti20}:
\begin{eqnarray}
\gamma_{\ell m} = - \sqrt{\frac{(\ell+2)(\ell - 1)}{\ell(\ell+1)}}\left( \kappa_{E,\ell m} + {\rm i} \kappa_{B,\ell m} \right)
\label{eq:KS}
\end{eqnarray}
In the above, $\kappa_{E}$ and $\kappa_{B}$ are the $E/B$ decomposition of the convergence maps, the latter being generally only second order and hence set to zero in this work. 
The inversion process is carried out with the polarised  {\sc map2alm} functions  in-built in {\sc Healpy}, and we further includes smoothing to suppress numerical noise, accomplished by convolving the $\kappa_E$ map with  a Gaussian beam with width $\sigma$=2.0 arcmin.

It is worth noting that this smoothing scale is relatively small, posing challenges for  the theoretical modelling of some probes within this regime.
For this reason, many analyses opt for modelling directly from simulations \citep[see][]{HD21, DESY3_Zuercher, HSCY1_peaks_sims}, bypassing some of the theoretical challenges.
%Note that this smoothing scale is quite small; some probes are difficult to model theoretically in that regime, in which case the modelling needs to be inferred from simulations directly \citep[as in][]{HD21, DESY3_Zuercher, HSCY1_Peaks_sims}.
In the interest of brevity, this paper does not cover statistics derived from aperture mass maps, though our methods are equally applicable to those estimators.
%We do not discuss statistics based on aperture mass maps in this paper for the sake of conciseness, but all our methods can be applied to those estimators as well.
Finally, while tomographic \gtwopcf data include auto-correlation and cross-redshift correlations, certain non-Gaussian statistics extract further information from analysing triplets, quadruplets, or quintuplets of redshift bins \citep{Martinet20}.
This work focuses only on auto-tomographic bins, yet our methods and results can be straightforwardly extended to include these higher-order redshift combinations.

@@

A number of higher-order statistics work at the level of shear catalogue,  and here we consider the following:
\begin{enumerate}
\item {\it Three-point correlation functions (or $M_{\rm ap}^3(\vartheta)$)}:  \JHD{Lucas/Laila, can you summarise the measurements and modelling, and provide references?}
\item {\it Squeezed bispectrum}: \JHD{ Anik, can you summarise the measurements and modelling, and provide references?}
\end{enumerate}

@@

In addition to the probes mentioned above, we further consider the following convergence map-based statistics:
\begin{enumerate}
    \item \textit{Convergence three-point correlation function}: \JHD{Alejandro, can you summarise the measurements and modelling, and provide references?}
    \item \textit{Peaks and Minima}: Local maxima and minima in the {\sc Healpix} convergence maps are counted in bins of signal-to-noise ratios, $\nu$, computed from the global  noise levels in the survey. Specifically, pure noise maps $\mathcal{N}$ are constructed from generating pure Gaussian maps with mean set to zero and variance set to $\sigma^2 =  \sigma_\epsilon^2 / (2 \Delta \Omega_{\rm pix} n_{\rm gal})$. Here, $\Delta \Omega_{\rm pix}$ is the average pixel area, $n_{\rm gal}$ is the mean galaxy density in our survey (0.6 gal / arcmin$^2$ per tomographic bin)  and $\sigma_\epsilon$ is the intrinsic dispersion in galaxy shapes, per component, which we set to 0.27.
   Peak count statistics are amongst the most widely used non-Gaussian lensing statistics, as it is simple to measure yet highly sensitive to non-Gaussian structures.
    Although some theoretical models have been developed to describe the largest peaks \citep[see][]{Shan18, HSCY1_Peaks_th}, we report here the results on a much wider range of statistics, which can only be accurately modelled from numerical simulations themselves.
    \item \textit{Lensing PDF}:  \JHD{Cora/Lina, can you summarise the measurements and modelling, and provide references?}
   % \item  \textit{Minkowski functionals}: \JHD{Nisha, can you summarise the measurements and modelling, and provide references?}
\end{enumerate}

These non-Gaussian statistics explore various physical scales and non-linear phenomena, leading to differences in their sensitivity to cosmology and systematic errors.
For instance, a statistical method effective in recovering cosmological information could be significantly influenced by secondary effects like intrinsic alignments (IA), casting doubts on its robustness.
Thus, incorporating IA  at the field level, with enough flexibility to account for the current uncertainty in our knowledge of the IA physics,  is essential for addressing these concerns (see Sec. \ref{sec:HOWLS}).
